---
title: "Data Types"
editor: visual
---

There are 6 types of data in R that are important to know, but the essential ones are logical, numeric, and character.

**Logical** Also known as boolean data, logical data is shown as TRUE or FALSE values:

```{r}
logical1 <- TRUE
logical2 <- FALSE

# The class() function outputs the data type of the object
class(logical1)


print(logical2)
print(class(logical2))

```

**Numeric** represents all data types that are real numbers with or with out decimal points.

```{r}

height <- 5.5
acres <- 1000

class(height)


print(acres)
print(class(acres))


```

**Character** specifies character or string values in a variable such as a singular character 'A' or a string of characters in 'Apple'.

```{r}

# Use '' or "" to show it's a string of characters
motorsports <- "formula1"

print(motorsports)
print(class(motorsports))


```

**Integer** specifies real values without decimal points

```{r}


# the suffix i specifies its imaginary data.
complex <- 3 + 2i

print(class(integer))

```

**Raw** specifies values as raw bytes

```{r}

# convert character to raw
raw_variable <- charToRaw("Welcome to Programiz")

print(raw_variable)
print(class(raw_variable))

# convert raw to character
char_variable <- rawToChar(raw_variable)

print(char_variable)
print(class(char_variable))

```
