---
title: "Histograms"
editor: visual
---

## Histograms

Frequency histograms are useful when you want to get an idea about the distribution of values in a numeric variable. The `hist()` is the function that plots histograms and takes a numeric vector as its main argument.

To set up this histogram, I construct a new variable called, `final_position`, and save it as a new object called `McLarenStandings_2023.`

```{r}

library(tidyverse)
library(RandomData)

McLarenStandings_2023 <- race_stats |>
  select(circuit, year, constructor, surname) |>
  # remove duplicates
  unique() |>
  filter(constructor == "McLaren" & year == 2023) |>
  mutate(
    final_position = case_when(
      #PIASTRI
      circuit == "Bahrain International Circuit" & surname == "Piastri" ~ "DNF",
      circuit == "Jeddah Corniche Circuit" & surname == "Piastri" ~ "15",
      circuit == "Albert Park Grand Prix Circuit" & surname == "Piastri" ~ "8", 
      circuit ==  "Baku City Circuit" & surname == "Piastri" ~ "11",
      circuit ==  "Miami International Autodrome" & surname == "Piastri" ~ "19",
      circuit == "Circuit de Monaco" & surname == "Piastri" ~ "10",
      circuit == "Circuit de Barcelona-Catalunya" & surname == "Piastri" ~ "13",
      circuit == "Circuit Gilles Villeneuve" & surname == "Piastri" ~ "11", 
      circuit == "Red Bull Ring" & surname == "Piastri" ~ "16",
      circuit == "Silverstone Circuit" & surname == "Piastri" ~ "4",
      circuit == "Hungaroring" & surname == "Piastri" ~ "5",
      circuit == "Circuit de Spa-Francorchamps" & surname == "Piastri" ~ "DNF",
      circuit == "Circuit Park Zandvoort" & surname == "Piastri" ~ "9",
      circuit == "Autodromo Nazionale di Monza" & surname == "Piastri" ~ "12",
      circuit == "Marina Bay Street Circuit" & surname == "Piastri" ~ "7",
      circuit == "Suzuka Circuit" & surname == "Piastri" ~ "3",
      circuit == "Losail International Circuit" & surname == "Piastri" ~ "2",
      circuit == "Circuit of the Americas" & surname == "Piastri" ~ "DNF",
      circuit == "Autódromo Hermanos Rodríguez" & surname == "Piastri" ~ "8",
      circuit == "Autódromo José Carlos Pace" ~ "14",
      circuit == "Las Vegas Strip Street Circuit" & surname == "Piastri" ~ "10",
      circuit == "Yas Marina Circuit" & surname == "Piastri" ~ "6",
        
        # NORRIS
        circuit == "Bahrain International Circuit" & surname == "Norris" ~ "17",
        circuit == "Jeddah Corniche Circuit" & surname == "Norris" ~ "17",
        circuit == "Albert Park Grand Prix Circuit" & surname == "Norris" ~ "6",
        circuit ==  "Baku City Circuit" & surname == "Norris" ~ "9",
        circuit ==  "Miami International Autodrome" & surname == "Norris" ~ "17", 
        circuit == "Circuit de Monaco" & surname == "Norris" ~ "9", 
        circuit == "Circuit de Barcelona-Catalunya" & surname == "Norris" ~ "17", 
        circuit == "Circuit Gilles Villeneuve" & surname == "Norris" ~ "13",
        circuit == "Red Bull Ring" & surname == "Norris" ~ "4", 
        circuit == "Silverstone Circuit" & surname == "Norris" ~ "2",
        circuit == "Hungaroring" & surname == "Norris" ~ "2",
        circuit == "Circuit de Spa-Francorchamps" & surname == "Norris" ~ "7",
        circuit == "Circuit Park Zandvoort" & surname == "Norris" ~ "9",
        circuit == "Autodromo Nazionale di Monza" & surname == "Norris" ~"8",
        circuit == "Marina Bay Street Circuit" & surname == "Norris" ~ "2",
        circuit == "Suzuka Circuit" & surname == "Norris" ~ "2",
        circuit == "Losail International Circuit" & surname == "Norris" ~ "3",
        circuit == "Circuit of the Americas" & surname == "Norris" ~ "3",
        circuit == "Autódromo Hermanos Rodríguez" & surname == "Norris" ~ "5",
        circuit == "Autódromo José Carlos Pace" ~ "2",
        circuit == "Las Vegas Strip Street Circuit" & surname == "Norris" ~ "DNF",
        circuit == "Yas Marina Circuit" & surname == "Norris" ~ "5"
        )
  ) 

# hist(data$x, 
#            col = "color", 
#            main = "Main Title",
#            xlab = "Label X-axsis")

```

In the following example, we will look at the distribution of the final positions of the Mclaren 2023 season for both of their drivers.

```{r}

# change from a character to numeric
McLarenStandings_2023 <- McLarenStandings_2023 |>
  mutate(final_position_numeric = ifelse(final_position == "DNF", 0, as.numeric(final_position)))

# make histogram
hist(McLarenStandings_2023$final_position_numeric, 
  # change color
  col = "orange", 
  # add main title
  main = "McLaren Race Results 2023", 
  # add title on x-axsis
  xlab = "Race Results")

```
