---
title: "Base R Descriptive Stats"
editor: visual
---

First, let's find the minimum speed recorded and the maximum speed recorded during the fastest laps for each race. Since fastest lap time is a character we need to change it from a character to a numeric value and lets remove any NAs.

```{r}
library(RandomData)
dat <- race_stats

# change NA to 0s and to numeric
dat$fastestLapSpeed <-as.numeric(
    ifelse(dat$fastestLapSpeed >= 0, dat$fastestLapSpeed, 0)
    )

```

-   Min and Max

    ```{r}
    min(dat$fastestLapSpeed)

    max(dat$fastestLapSpeed)

    range(dat$fastestLapSpeed)
    ```

-   Mean

    -   The **mean** is average value of all the numbers in a set.

```{r}

mean(dat$fastestLapSpeed)

```

-   Median

    -   The **median** is the middle value in a set of numbers when they are ordered from least to greatest.

```{r}

median(dat$fastestLapSpeed)

```

-   First and Third Quartiles

    -   The **first quartile** range is the value under which 25 percent of the data points are found when they are arranged in increasing order, and the **third quartile** range is where 75 percent of the data points are found when they are arranged in increasing order

```{r}

quantile(dat$fastestLapSpeed, 0.25)

quantile(dat$fastestLapSpeed, 0.75)

```

-   IQR

    -   The **IQR** is the difference between the first and third quartile.

```{r}

IQR(dat$fastestLapSpeed)

```

-   Standard Deviation and Variance

    -   **Variance** is the average squared difference between data points in a set, which measures how much the values in a set vary from each other. While **Standard Deviation** is the measure of how far the values in a set are from the mean

```{r}

sd(dat$fastestLapSpeed)

var(dat$fastestLapSpeed)

```
