---
title: "Scatter Plot"
editor: visual
---

## Scatter Plots

The first plot style we will be looking at is a scatter plot. To plot a scatter plot in `ggplot` you will use the format discussed above and to tell r you want it to graph a scatter plot you will use the geom function; `geom_point()`.

```{r}
#| include: false

library(tidyverse)
library(RandomData)

# ggplot(data, aes(x, y) + geom_point()

```

In the following example, we will plot the qualifying position for the 20 drivers and the final points they were awarded for the Las Vegas Grand Prix in 2023.

```{r}

scatterplot <- race_stats |>
  filter(circuit == "Las Vegas Strip Street Circuit" & year == "2023")

ggplot(scatterplot, aes(quali_position, points)) + 
        geom_point()

```

Let's add a main title and adjust the x and y labels.

```{r}
ggplot(scatterplot, aes(quali_position, points)) + 
        geom_point() +
        labs(
          x = "Qualifying Position",
          y = "Points Awarded",
          title = "Qualifying Position vs. Points Awarded at Las Vegas GP 2023"
        )
```

If we wanted to add a smoother line we would use the `geom_smooth()` and we will be fitting a linear regression line, thus inside `geom_smooth()` we `method = "lm"`.

```{r}

ggplot(scatterplot, aes(quali_position, points)) + 
        geom_point() +
        labs(
          x = "Qualifying Position",
          y = "Points Awarded",
          title = "Qualifying Position vs. Points Awarded at Las Vegas GP 2023"
        ) +
        geom_smooth(method = "lm")

```
