---
title: "Time Series"
editor: visual
---

## Time Series

`geom_line()` in `ggplot2` is used to create a line plot by connecting data points with a continuous line, which is ideal for visualizing trends over time. It is particularly useful for time series data because it clearly shows how a variable changes across ordered time intervals, allowing for easy identification of patterns and trends.

```{r}

#| include: false
library(tidyverse)
library(RandomData)

```

```{r}

?constructors_stats

# Calculate total wins per year for Ferrari
ferrari_wins <- constructors_stats |>
  filter(constructor == "Ferrari") |>
  group_by(year) |> 
  summarize(total_wins = sum(max(constructor_wins)))

ggplot(ferrari_wins, aes(x = year, y = total_wins)) +
  geom_line(color = "red", linewidth = 1.2) +  # Creates a red line plot
  labs(title = "Ferrari Wins 1958 to 2024 (Before the Summer Break)", 
       x = "Year", 
       y = "Wins") +
  theme_classic()

```

-   **`aes(x = year, y = total_wins)`**:

    -   `x = year`: Puts the years on the x-axis.

    -   `y = total_wins`: Puts the total number of wins on the y-axis.

-   **`geom_line(color = "red", linewidth = 1.2)`**:

    -   `geom_line()`: Creates a **line plot** instead of points.

    -   `color = "red"`: Adds a color to the line

    -   `linewidth = 1.2`: Makes the line slightly thicker for better visibility.
