---
title: "Bar Plots"
---

## Bar Plot

Now, lets say we wanted to compare this across different categories, therefore instead we can use bar plots instead of histograms. In the following example, we will look at the distribution of the final positions of the McLaren 2023 season for both of their drivers. To set up this histogram, I construct a new variable called, `final_position`, and save it as a new object called `McLarenStandings_2023.`

```{r}
#| include: false
library(tidyverse)
library(RandomData)
```

```{r}

McLarenStandings_2023 <- race_stats |>
  select(circuit, year, constructor, surname) |>
  # remove duplicates
  unique() |>
  filter(constructor == "McLaren" & year == 2023) |>
  mutate(
    final_position = case_when(
      #PIASTRI
      circuit == "Bahrain International Circuit" & surname == "Piastri" ~ "DNF",
      circuit == "Jeddah Corniche Circuit" & surname == "Piastri" ~ "15",
      circuit == "Albert Park Grand Prix Circuit" & surname == "Piastri" ~ "8", 
      circuit ==  "Baku City Circuit" & surname == "Piastri" ~ "11",
      circuit ==  "Miami International Autodrome" & surname == "Piastri" ~ "19",
      circuit == "Circuit de Monaco" & surname == "Piastri" ~ "10",
      circuit == "Circuit de Barcelona-Catalunya" & surname == "Piastri" ~ "13",
      circuit == "Circuit Gilles Villeneuve" & surname == "Piastri" ~ "11", 
      circuit == "Red Bull Ring" & surname == "Piastri" ~ "16",
      circuit == "Silverstone Circuit" & surname == "Piastri" ~ "4",
      circuit == "Hungaroring" & surname == "Piastri" ~ "5",
      circuit == "Circuit de Spa-Francorchamps" & surname == "Piastri" ~ "DNF",
      circuit == "Circuit Park Zandvoort" & surname == "Piastri" ~ "9",
      circuit == "Autodromo Nazionale di Monza" & surname == "Piastri" ~ "12",
      circuit == "Marina Bay Street Circuit" & surname == "Piastri" ~ "7",
      circuit == "Suzuka Circuit" & surname == "Piastri" ~ "3",
      circuit == "Losail International Circuit" & surname == "Piastri" ~ "2",
      circuit == "Circuit of the Americas" & surname == "Piastri" ~ "DNF",
      circuit == "Autódromo Hermanos Rodríguez" & surname == "Piastri" ~ "8",
      circuit == "Autódromo José Carlos Pace" ~ "14",
      circuit == "Las Vegas Strip Street Circuit" & surname == "Piastri" ~ "10",
      circuit == "Yas Marina Circuit" & surname == "Piastri" ~ "6",
        
        # NORRIS
        circuit == "Bahrain International Circuit" & surname == "Norris" ~ "17",
        circuit == "Jeddah Corniche Circuit" & surname == "Norris" ~ "17",
        circuit == "Albert Park Grand Prix Circuit" & surname == "Norris" ~ "6",
        circuit ==  "Baku City Circuit" & surname == "Norris" ~ "9",
        circuit ==  "Miami International Autodrome" & surname == "Norris" ~ "17", 
        circuit == "Circuit de Monaco" & surname == "Norris" ~ "9", 
        circuit == "Circuit de Barcelona-Catalunya" & surname == "Norris" ~ "17", 
        circuit == "Circuit Gilles Villeneuve" & surname == "Norris" ~ "13",
        circuit == "Red Bull Ring" & surname == "Norris" ~ "4", 
        circuit == "Silverstone Circuit" & surname == "Norris" ~ "2",
        circuit == "Hungaroring" & surname == "Norris" ~ "2",
        circuit == "Circuit de Spa-Francorchamps" & surname == "Norris" ~ "7",
        circuit == "Circuit Park Zandvoort" & surname == "Norris" ~ "9",
        circuit == "Autodromo Nazionale di Monza" & surname == "Norris" ~"8",
        circuit == "Marina Bay Street Circuit" & surname == "Norris" ~ "2",
        circuit == "Suzuka Circuit" & surname == "Norris" ~ "2",
        circuit == "Losail International Circuit" & surname == "Norris" ~ "3",
        circuit == "Circuit of the Americas" & surname == "Norris" ~ "3",
        circuit == "Autódromo Hermanos Rodríguez" & surname == "Norris" ~ "5",
        circuit == "Autódromo José Carlos Pace" ~ "2",
        circuit == "Las Vegas Strip Street Circuit" & surname == "Norris" ~ "DNF",
        circuit == "Yas Marina Circuit" & surname == "Norris" ~ "5"
        )
  ) |> mutate(final_position_numeric = as.numeric(final_position))

# Summarize mean race positions by driver
barplot <- McLarenStandings_2023 |>
  group_by(surname) |>
  summarize(mean_position = mean(final_position_numeric))

# Create a bar plot of mean race positions
ggplot(barplot, aes(x = mean_position, y = surname, fill = surname)) +
  geom_col() +
  labs(title = "McLaren Mean Race Results 2023", 
       x = "Mean Race Position", 
       y = "McLaren Driver") +
  scale_fill_manual(values = c("orange", "grey")) + 
  theme_minimal()
```

-   **`aes(x = mean_position, y = surname, fill = surname)`**:

    -   `x = mean_position`: The x-axis represents the **mean race position** of each driver.

    -   `y = surname`: The y-axis represents the **driver's name**.

    -   `fill = surname`: Each driver gets a unique color for their bar.

-   **`geom_col()`**:

    -   Creates a **bar chart** where the bar length represents the **mean race position**

-   **`scale_fill_manual(values = c("orange", "grey"))`**:

    -   Manually assigns colors to the bars.

    -   `"orange"` for one driver and `"grey"` for the other (representing McLaren's team colors).
