---
title: "Times Series"
editor: visual
---

## Time Series

A time series plot looks at a variable over time to see trends. To make a time series graph, we use the same function as scatter plots, `plot()`, however, we change the type of graph from a point to a line graph. We can do this by changing the `type =` in the function from a `type = "p"` to a `type = "l"`. In addition, the variable on the x-axis should always be a variable that measures time.

```{r}

library(tidyverse)
library(RandomData)

# Time Series Plot
# plot(data$y ~ data$timevariable,
#                type ="l",
#                col = "color", 
#               main = "Main Title",
#               ylab = "Label Y-axsis",
#               xlab = "Label X-axsis")

```

In the following example, we look at the total wins for Ferrari drivers over time.

```{r}

# Calculate total wins per year for Ferrari
ferrari_wins <- constructors_stats |>
  filter(constructor == "Ferrari") |>
  group_by(year) |> 
  summarize(total_wins = sum(max(constructor_wins)))

## Make Time Series plot
plot(ferrari_wins$total_wins ~ ferrari_wins$year,
     # make it a line not points
     type = "l", 
     # add color
     col = c("red"), 
     # change width of the line
     lwd = 2, 
     # add main title
     main = "Ferrari Wins 1958 to 2024 (Before the Summer Break)",
      # add title on x-axsis
     xlab = "Year", 
      # add title on y-axsis
     ylab = "Wins")
```

## 
