---
title: "Setting up a Working Directory"
---

It is important to keep all the files associated with a given project, such as data, scripts, analytical results, and figures, together. To do this we need to make a working directory.

### To create a new working directory in RStudio:

1.  Use **File → New Project** or use the **New Project** button (available on the Projects toolbar in the top right corner or on the global toolbar at the top left corner).
2.  This will open the “New Project Wizard” popup.
3.  Click "New Directory".

### To set the working directory, there are two possible ways, using code or using RStudio directly:

#### Option 1: Using R Code

```{r}
#| eval: false
#| include: true

# Set working directory (replace this with your own)
setwd("/Users/yourname/Desktop/filename/filename")

```

#### Option 2: Using RStudio Interface

1.  On the top of the RStudio menu, click on **Session**.

2.  Click **Set Working Directory**.

3.  Then click **To Source File Location**.

To check if the working directory is set correctly, use the following code:

```{r}

#| eval: false
#| include: true

# Returns the filepath of the current working directory
getwd()

```
